\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname}

This dissertation would not be possible without the help and support of many people. First of all, I would like to thank my supervisor, Prof. Alex Biryukov, for leading me through my doctoral studies and especially for giving me freedom in the choice of research topics. I would like to thank my thesis supervision committee members, Prof. Jean-Sébastien Coron, and Prof. Volker Müller, for following my research and giving valuable advice. I am grateful to Prof. Gregor Leander and to Dr. Matthieu Rivain for agreeing to serve on my defense as jury members.

I would like to thank the Luxembourg National Research Fund for funding my research (project reference 9037104) and Pierre Fuhrer for coordinating the project. I am grateful for the hospitable country of Luxembourg, the University of Luxembourg and to the Interdisciplinary Centre for Security, Reliability and Trust for providing me an excellent research environment.

It was very exciting and educative for me to work on interesting topics with my coauthors, which include 
Alex Biryukov,
Christof Beierle,
Daniel Dinu,
Johann Gro{\ss}sch{\"a}dl,
L\'eo~Perrin,
Luan Cardoso dos Santos,
Qingju Wang,
Vesselin Velichkov,
Yann Le Corre.

I am thankful for all colleagues that I have met at the university for interesting  discussions by the cup of coffee:
% Alex Biryukov,
Benoît Cogliati,
Brian Shaft,
% Christof Beierle,
Dag Arne Osvik,
% Daniel Dinu,
Daniel Feher,
Dmitry Khovratovich,
Giuseppe Vito,
Ivan Pustogarov,
% Johann Gro{\ss}sch{\"a}dl,
% Luan Cardoso Dos Santos,
% Léo Perrin,
Moon Sung Lee,
Najmeh Soroush,
Patrick Derbez,
Praveen Vadnala,
% Qingju Wang,
Ritam Baumik,
Sergei Tikhomirov,
Shange Fu,
Srinivas Vivek,
% Vesselin Velichkov,
Vitor Pereira.
% Yann Le Corre.
I am also delighted to have met interesting people from other research groups:
Antonio Di Maio, Balazs Pejo, Christoph Lambert, Ivana Vukotic, Chista Nadimi, Masoud Tabatabaei, Marharyta Aleksandrova, Vincent Rahli.

I am grateful to the university secretaries and to the doctoral school for their administrative support: Fabienne Schmitz, Claudia Thür, Natali Kirf, Jessica Giro, Céline Lecarpentier.

My warmest gratitude is to my family for their support, especially to my mother, to my wife Asya and to my daughter Nicole.

I would like to acknowledge organizers of various competitions that I participated in: the \href{http://whibox.cr.yp.to/}{WhibOx} Contest, the \href{https://sites.google.com/site/skinnycipher/cryptanalysis-competition/2018-2019-competition}{SKINNY} cryptanalysis competition, the \href{https://nsucrypto.nsu.ru/}{NSUCRYPTO} Olympiad and all the \href{https://ctftime.org}{Capture-The-Flag} competitions that I participated as a part of the teams \href{https://ctftime.org/team/1005}{MSLC} and \href{https://ctftime.org/team/15726}{LC{\Lightning}BC}. Competitions always kept me inspired and willing to learn new things.

Finally, I appreciate the work of the developers of the open source software that I used, including the Linux Ecosystem, \LaTeX, Python and SageMath~\cite{sage}, TikZ for Cryptographers \cite{TikZ}, Git.
\end{acknowledgements}